Embedding Intuitionistic into Classical Logic

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
space reduction, updated conclusion

\textbf{Related Work.}
Several approaches to obtain constructive content from classical proofs have been studied in the literature.
An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
Further circumstances under which this entailment holds have been identified in~\cite{schwichtenberg}, however this approach cannot encapsulate the general case.\todo{can you elaborate a few words on ``further circumstances'' and ``the general case''}
Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
In this case we necessarily lose realizability at least of $\exists$ and $\vee$, which --- one could argue --- is why we want a program in the first place.\todo{as the curry-howard correspondence is only mentioned in passing now, we should adapt the last part of sentence}

@@

\begin{definition}
A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
	\end{itemize}
We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid:
\begin{example}\label{LEMcounterexample}
Consider the famous law of excluded middle $A\vee\neg A$. Recall that $\neg A := A\to \bot$. As a counter-model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$. Then $u\not\models A$ and $u\not\models \neg A$, i.e. $u\not\models A\vee\neg A$.
\end{example}
In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable. This is insufficient for intuitionistic logic, e.g. $A\vee\neg A$ is not an intuitionistic tautology but its negation is not satisfiable either.

@@

\begin{definition}
Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
		m, &\text{if $b=a$,}\\
		v(b), &\text{otherwise.}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$ and  $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. A formula is valid if every $\Sigma$-structure satisfies it. We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
\end{definition}

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
	\begin{itemize}
		\item $M_u\subseteq M_w$.
		\item $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
		\item $R^{I_w}|_{M_u} = R^{I_u}$.
	\end{itemize}
	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
	\begin{itemize}
		\item $u, v\not\models\bot$ and $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$.
		\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$.
		\item $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. A formula is valid if it is satisfied in eveery $\Sigma$-Kripke structure, it is valid for some Frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that Frame. We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
Analogous remarks with regard to validity and satisfiability hold as in the propositional case.

@@

\section{Equivalid normalization}
\label{section:tseytin}

\todo{find better name for section, as the normalization works for equisatsifiability as well as for equivalidity}

@@

We give a transformation of structures that corresponds to the above.

\begin{definition}
	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$. By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables. In the classical case have
	$$\begin{matrix}
		P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\indent P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}
	\end{matrix}$$
	and for intuitionistic logic and each world $u$
	$$\begin{matrix}
		P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\indent P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
		 P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
		 P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}
	\end{matrix}$$
\end{definition}

The above definitions are motivated by the following results which follow directly from the involved definitions and are true both for classical and intuitionistic semantics:

\begin{lemma}
	For every formula $\varphi$ and structure $\mathcal M$
	$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$
\end{lemma}

@@

\begin{definition}
Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
We say $\psi\in\mathcal F_\to$ is \emph{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true. For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
		u,&\text{ if $\psi$ is fulfilled in $u$,}\\
		f^I_\psi(u),&\text{ else.}		
	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
		\item Interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $\psi_1\dots\psi_n$,}\\
			\psi_1\dots\psi_n\psi, &\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$, we set $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)).$$
	\end{itemize}
\end{definition}
\begin{lemma}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
	\end{enumerate}

\end{lemma}

\begin{proof}

	1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.

Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

@@

An examples of this translation can be found in the appendix.

@@

So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
	x, &\text{ if $\psi$ occurs in $x$,}\\
	x, \psi&\text{ otherwise.}
\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula. This gives us the translation stated in Theorem~\ref{thm:prop-translation}.
\begin{theorem}
	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
\end{theorem}

@@

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

@@

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
\begin{proof}
Suppose $\varphi$ is not intuitionistically valid i.e. it $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then by the precious section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$. Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.

@@

As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
\begin{itemize}
	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
	\item World($u$) encoding that $u$ represents a world.
	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
\end{itemize}
Then analogously to the predicate case define
\begin{align*}
	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
				 & \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
				 & \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
\end{align*}
Example encodings could be
\begin{align*}
	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
	\text{World}(u) = &\:\exists xE(x, u)\\
	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
\end{align*}

@@

Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.

@@

\begin{definition}
\label{def:fo-fulfilled}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
\end{definition}

@@

%We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
\todo{can you add a few sentences on proof translations, which is an important goal because it allows program extraction using the curry-howard correspondence?}

@@

\section{Omitted Proofs}

\todo{the lemmata still use the set `T', so the notation is probably not consistent anymore}

>>>>

\textbf{Related Work.}
Several approaches to obtain constructive content from classical proofs have been studied in the literature.
An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
Other very restricted subsets of classical logic have been identified under which this entailment holds, e.g. in~\cite{schwichtenberg}, however this approach is not applicable in general.
Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
In this case we necessarily lose realizability at least of $\exists$ and $\vee$ in general.

@@

\begin{definition}
A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
	\end{itemize}
We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
		m, &\text{if $b=a$,}\\
		v(b), &\text{otherwise.}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$, $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$, $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. A formula is valid if every $\Sigma$-structure satisfies it. We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
\end{definition}

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
	\begin{itemize}
		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. A formula is valid if it is satisfied in eveery $\Sigma$-Kripke structure, it is valid for some Frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that Frame. We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
Analogous remarks with regard to validity and satisfiability hold as in the propositional case.

@@

\section{Normal form translation}
\label{section:tseytin}

@@

We give a transformation of structures that corresponds to the above, i.e. for each structure $\mathcal M$ and formula $\varphi$ wee can give a structure $\mathcal S(\mathcal M, \varphi)$, the exact definition of which can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
\begin{lemma}
	For every formula $\varphi$ and structure $\mathcal M$
	$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$
\end{lemma}

@@

\begin{definition}
Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
We say $\psi\in\mathcal F_\to$ is \emph{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true. For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
		u,&\text{ if $\psi$ is fulfilled in $u$,}\\
		f^I_\psi(u),&\text{ else.}		
	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
		\item Interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $\psi_1\dots\psi_n$,}\\
			\psi_1\dots\psi_n\psi, &\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$, we set $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)).$$
	\end{itemize}
\end{definition}
\begin{lemma}~\label{thm:prop-countermodel-reduction}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
	\end{enumerate}
\end{lemma}

A proof can be found in the appendix as lemma~\ref{thm:prop-countermodel-reduction2}. Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

@@

An example of this translation can be found in the appendix as Example~\ref{ex:prop-tree-model}.

@@

So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
	x, &\text{ if $\psi$ occurs in $x$,}\\
	x, \psi&\text{ otherwise.}
\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula. This gives us the translation stated in Theorem~\ref{thm:reduction-propositional}.
\begin{theorem}
	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
\end{theorem}

@@

Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

@@

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
A proof can be found in the appendix as proof~\ref{lemma:QBF}.

@@

As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
\begin{itemize}
	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
	\item World($u$) encoding that $u$ represents a world.
	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
\end{itemize}
Then analogously to the predicate case define
\begin{align*}
	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
\end{align*}
Example encodings could be
\begin{align*}
	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
	\text{World}(u) = &\:\exists xE(x, u)\\
	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
\end{align*}

@@

Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced. A proof is given in the appendix as Lemma~\ref{lemma:fo-simplification}.

@@

\begin{definition}
\label{def:fo-fulfilled}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
\end{definition}

We get an analogous result to lemma~\ref{thm:prop-countermodel-reduction}.

@@

As a next step we will benchmark our translation using state-of-the art automated theorem provers, in particular the Vampire theorem prover~\cite{kovacs2013first}. We will have to adapt the counter-model translation to a proof translation for the particular calculus, e.g. superposition in the case of Vampire. Ultimately we hope that this will also open a new path for program extraction via the Curry-Howard Correspondence.

@@

\section{Omitted content}

\begin{lemma}\label{ap1}
	Let $\chi$ be a formula and $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
	\begin{itemize}
		\item if $\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
	\end{itemize}
\end{lemma}
